Outbreaks in Toronto Healthcare Institutions After COVID-19

### 2024-09-26 paper/paper.qmd
update required graphs/plots in the analysis

---
title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
subtitle: "Analysis of the Frequency Trends and the Causative Agent"
author: 
  - Yunkyung Ko
thanks: "Code and data are available at: https://github.com/koyunkyung/paper1_draft.git."
date: today
date-format: long
abstract: "In Canada, outbreak incidence has been slowly increasing from June 2024 up to the beginning of the accumulating data period on August 10, 2024. (ADD CITATION https://health-infobase.canada.ca/covid-19/) Using the 'Outbreaks in Toronto Healthcare Institutions' dataset, we show how the trends of outbreak occurrence has been over the 5 years after COVID-19. (specify headline result). (implications)."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: fig-outbreak
#| fig-cap: Total Outbreaks Count per 6-month
#| echo: false

data |>
  ggplot(mapping = aes(x = date_of_outbreak)) +
  geom_bar() +
  theme_minimal()+
  labs(x = "Date Period", y = "Number of Outbreaks Reported")

@@

@@

To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
```{r}
# Process the data to get the top 3 causative agents, excluding 'NA' and 'Unable to identify'
top_causative_agents <- cleaned_data |>
  janitor::clean_names() |>

  # Filter out rows where causative_agent is NA or 'Unable to identify'
  filter(!is.na(causative_agent) & causative_agent != "Unable to identify") |>

  # Group by causative agent and count occurrences
  group_by(causative_agent) |>
  summarise(total_cases = n()) |>

  # Arrange in descending order of total cases
  arrange(desc(total_cases)) |>

  # Select the top 3 causative agents
  slice_max(order_by = total_cases, n = 3)

# View the top 3 causative agents
top_causative_agents
```

```{r}
#| label: fig-causations
#| fig-cap: Outbreaks Count per 6-month, by Top 3 Causative Agent
#| echo: false
#| warning: false
#| message: false

### NEED A LINE PLOT WITH THREE LINES, EACH INDICATING THE TOP 3 CAUSATIVE AGENTS.

@@

Talk way more about it.

>>>>

---
title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
subtitle: "Analysis of the Frequency Trends and the Causative Agent"
author: 
  - Yunkyung Ko
thanks: "Code and data are available at: https://github.com/koyunkyung/toronto_outbreaks"
date: today
date-format: long
abstract: "In Canada, outbreak incidence has been slowly increasing from June 2024 up to the beginning of the accumulating data period on August 10, 2024. (ADD CITATION https://health-infobase.canada.ca/covid-19/) Using the 'Outbreaks in Toronto Healthcare Institutions' dataset, we show how the trends of outbreak occurrence has been over the 5 years after COVID-19. (specify headline result). (implications)."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}

@@

```{r}
#| label: fig-outbreak
#| fig-cap: Total Outbreaks Count in Toronto Healthcare Institutions
#| echo: false

data |>
  ggplot(mapping = aes(x = year)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Date Period", y = "Number of Outbreaks Reported",
       title = "Total Number of Outbreaks Reported in Toronto Healthcare Institutions") +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))

```{r}
#| label: fig-setting
#| fig-cap: Outbreaks Count by Outbreak Settings
#| echo: false

library(dplyr)
library(ggplot2)

outbreak_counts <- data %>%
  count(outbreak_setting)

ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) + 
  theme_minimal() +
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
       title = "Number of Outbreaks Reported by Setting") +
  theme(axis.text.x = element_blank(),  
        axis.ticks = element_blank(),  
        panel.grid = element_blank())  
```

@@

# Result

To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
```{r}

@@

```{r}
#| label: fig-outbreaks count by types
#| fig-cap: Outbreaks Count in Toronto Healthcare Institutions, by Types of Outbreak
#| echo: false
#| warning: false
#| message: false

# Summarizing the data to get the number of outbreaks per month and type of outbreak
annual_data <- data %>%
  group_by(year, type_of_outbreak) %>%
  summarise(outbreak_count = n()) %>%
  ungroup()

# Convert the year column to a factor with levels in chronological order
annual_data$year <- factor(annual_data$year, levels = unique(annual_data$year))

# Create the line plot
ggplot(annual_data, aes(x = year, y = outbreak_count, group = type_of_outbreak, color = type_of_outbreak)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
  labs(title = "Trend of Outbreak Counts by Type (2016-2024)",
       x = "Year", y = "Outbreak Count", color = "Type of Outbreak") +
  theme_minimal() +
  scale_x_discrete(drop = FALSE) # Ensure all years are shown on the x-axis

```{r}
#| label: fig-respiratory outbreaks count by outbreak settings
#| fig-cap: Respiratory Outbreaks Count, by Outbreak Settings
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)

# Filter for 'Respiratory' outbreaks
respiratory_data <- data %>%
  filter(type_of_outbreak == "Respiratory")

# Aggregate data by year and outbreak setting
annual_counts <- respiratory_data %>%
  group_by(year, outbreak_setting) %>%
  summarise(outbreak_count = n(), .groups = 'drop')

# Create a stacked bar plot
ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
  geom_col(position = "stack") +  # Stack bars based on outbreak settings
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  labs(title = "Annual Respiratory Outbreak Counts by Setting",
       x = "Year",
       y = "Outbreak Count",
       fill = "Outbreak Setting") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
```

```{r}
#| label: fig-enteric outbreaks count by outbreak settings
#| fig-cap: Enteric Outbreaks Count, by Outbreak Settings
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)

# Filter for 'Enteric' outbreaks
enteric_data <- data %>%
  filter(type_of_outbreak == "Enteric")

# Aggregate data by year and outbreak setting
annual_counts <- enteric_data %>%
  group_by(year, outbreak_setting) %>%
  summarise(outbreak_count = n(), .groups = 'drop')

# Create a stacked bar plot
ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
  geom_col(position = "stack") +  # Stack bars based on outbreak settings
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  labs(title = "Annual Enteric Outbreak Counts by Setting",
       x = "Year",
       y = "Outbreak Count",
       fill = "Outbreak Setting") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
```
Talk way more about it.
